---
title: "Data generation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data generation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 10,
  fig.height = 6,
  fig.align = "center"
)
```

This vignette shows how to build stochastic models, combine them with `+`,
generate data, and plot the results.

```{r}
library(gmwmx2)
set.seed(123)
```

# Available models

You can build a single stochastic process with any of the following model
constructors:

- `wn()` (white noise)
- `ar1()` (AR(1))
- `pl()` (power-law)
- `matern()` (Matérn)
- `rw()` (random walk)
- `flicker()` (flicker / 1/f)

Each constructor returns a `time_series_model` object that can be plotted or
combined with others using `+`.

# Single model generation and plotting

```{r single-model-plot, fig.alt = "Simulated AR(1) time series."}
model_single <- ar1(phi = 0.9, sigma2 = 1)
series_single <- generate(model_single, n = 500)
plot(series_single)
```

The generated object is a `generated_time_series` with a numeric series in
`$series`:

```{r}
head(series_single$series)
```

# Generate and plot each base model

```{r all-base-models, fig.alt = "Simulated time series from each base stochastic model."}
base_models <- list(
  wn(sigma2 = 1),
  ar1(phi = 0.7, sigma2 = 1),
  pl(kappa = 0.3, sigma2 = 2),
  matern(sigma2 = 1, lambda = 0.5, alpha = 1.0),
  rw(sigma2 = 0.2),
  flicker(sigma2 = 1)
)

par(mfrow = c(3, 2), mar = c(3, 4, 2, 1))
for (m in base_models) {
  s <- generate(m, n = 500)
  plot(s)
}
par(mfrow = c(1, 1))
```

# Composite models (sum of processes)

Use `+` to build a composite model from multiple stochastic processes. The
result is a `sum_model` that can be passed to `generate()`.

```{r composite-model-plot, fig.alt = "Simulated composite time series with white noise, AR(1), and power-law components."}
model_comp <- wn(sigma2 = 1) + ar1(phi = 0.8, sigma2 = 0.5) + pl(kappa = 0.3, sigma2 = 2)
series_comp <- generate(model_comp, n = 500)
plot(series_comp)
```

# More composite model examples

```{r composite-models-plot, fig.alt = "Simulated time series from multiple composite stochastic models."}
comp_models <- list(
  wn(sigma2 = 2) + rw(sigma2 = 0.1),
  wn(sigma2 = 1) + ar1(phi = 0.95, sigma2 = 0.3) + flicker(sigma2 = 0.5),
  pl(kappa = 0.4, sigma2 = 1) + matern(sigma2 = 0.8, lambda = 0.6, alpha = 1.2)
)

par(mfrow = c(3, 1), mar = c(3, 4, 2, 1))
for (m in comp_models) {
  s <- generate(m, n = 500)
  plot(s)
}
par(mfrow = c(1, 1))
```

The composite output is a `generated_composite_model_time_series` with:

- `series` (the total sum)
- `components` (a list of each component series)
- `model` (names of each component)

```{r}
names(series_comp)
```

# Combining and plotting

The `plot()` method for a composite model stacks each component vertically and
adds the sum at the top. For single models, `plot()` draws the time series with
the model name in the title.
