---
title: "Estimate composite stochastic processes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Estimate composite stochastic processes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 10,
  fig.height = 6,
  fig.align = "center"
)
```

This vignette shows how to estimate composite stochastic models using
`gmwm2()`. We generate data from known models, fit composite candidates,
and visualize the results.

We consider a zero-mean stochastic process \(\{Y_t\}_{t=1}^n\) generated from a
composite model parameterized by \(\boldsymbol{\gamma} \in \boldsymbol{\Gamma}\).

Given a parametric model with parameters \(\boldsymbol{\gamma}\), the GMWM
estimator computes $\boldsymbol{\gamma}$ by solving:
\begin{equation}
    \hat{\boldsymbol{\gamma}} = \underset{\boldsymbol{\gamma} \in \boldsymbol{\Gamma}}{\arg\min} \left\{\hat{\boldsymbol{\nu}} - \boldsymbol{\nu}(\boldsymbol{\gamma})\right\}^T \boldsymbol{\Omega} \left\{\hat{\boldsymbol{\nu}} - \boldsymbol{\nu}(\boldsymbol{\gamma})\right\},
\end{equation}

where \(\hat{\boldsymbol{\nu}}\) is the empirical wavelet variance of the
observed series and \(\boldsymbol{\nu}(\boldsymbol{\gamma})\) is the theoretical
wavelet variance implied by the model.


```{r}
library(gmwmx2)
set.seed(123)
```


```{r}
n <- 10000
mod1 <- wn(sigma2 = 25) + ar1(phi = 0.99, sigma2 = 1)
y1 <- generate(mod1, n = n)
plot(y1)
fit1 <- gmwm2(y1, model = wn() + ar1())
fit1
plot(fit1)
```

```{r}
mod2 <- wn(5) + pl(kappa = -0.9, sigma2 = 1)
y2 <- generate(mod2, n = n)
plot(y2)
fit2 <- gmwm2(y2, model = wn() + pl())
fit2
plot(fit2)
```

```{r}
mod3 <- wn(5) + ar1(phi = 0.98, sigma2 = 1) + rw(0.1)
y3 <- generate(mod3, n = n)
plot(y3)
fit3 <- gmwm2(y3, model = wn() + ar1() + rw())
fit3        
plot(fit3)
```


```{r}
mod4 <- wn(20) + rw(0.1)
y4 <- generate(mod4, n = n)
plot(y4)
fit4 <- gmwm2(y4, model = wn() + rw())
fit4
plot(fit4)
```
