---
title: "Estimate geodetic time series with dependent errors and missing observations"
output: rmarkdown::html_vignette
bibliography: REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{Estimate geodetic time series with dependent errors and missing observations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 10, # Set default plot width (adjust as needed)
  fig.height = 8, # Set default plot height (adjust as needed)
  fig.align = "center" # Center align all plots
)

# knitr::opts_chunk$set(eval = FALSE)
```

# Model and estimator
The `gmwmx2` R package supports estimation of linear models with dependent
errors described by a composite stochastic model in the presence of missing
data.

More precisely, we assume the following model:
\begin{equation}
    \boldsymbol{Y} = \boldsymbol{X} \boldsymbol{\beta} + \boldsymbol{\varepsilon}, \quad \boldsymbol{\varepsilon} \sim \mathcal{F}\left\{\boldsymbol{\Sigma}\left(\boldsymbol{\gamma}\right)\right\},
\end{equation}

where $\boldsymbol{X} \in \mathbb{R}^{n \times p}$ is a design matrix of
observed predictors, $\boldsymbol{\beta} \in \mathbb{R}^p$ is the regression
parameter vector, and $\boldsymbol{\varepsilon} = \{\varepsilon_{i}\}_{i=1,\ldots,n}$
is a zero-mean process following an unspecified joint distribution
$\mathcal{F}$ with positive-definite covariance function
$\boldsymbol{\Sigma}(\boldsymbol{\gamma}) \in \mathbb{R}^{n\times n}$,
characterizing the second-order dependence structure of the process and
parameterized by the vector $\boldsymbol{\gamma} \in \boldsymbol{\Gamma} \subset \mathbb{R}^q$. 

We then define a random variable $\boldsymbol{Z} = \{Z_{i}\}_{i=1,\ldots,n}$
that describes the missing-observation mechanism. More specifically, the vector
$\boldsymbol{Z} \in \{0, 1\}^n$ is a binary-valued stationary process
independent of $\boldsymbol{Y}$ with expectation
$\mu(\boldsymbol{\vartheta}) = \mathbb{E}[Z_i] \in [0, 1)$, for all $i$, and
covariance matrix $\boldsymbol{\Lambda}(\boldsymbol{\vartheta}) = \mathbb{V}[\boldsymbol{Z}]
\in \mathbb{R}^{n\times n}$ whose structure is assumed known up to the
parameter vector $\boldsymbol{\vartheta} \in \boldsymbol{\Upsilon} \subset \mathbb{R}^k$.


We observe only the stochastic process $\tilde{\boldsymbol{Y}} = \boldsymbol{Z} \odot \boldsymbol{Y}$,
where $\odot$ denotes the Hadamard product. Hence,
$\tilde{\boldsymbol{Y}} \in \mathbb{R}^n$ represents the observed process
vector with null elements at positions where observations are missing. 

Using $\otimes$ to denote the Kronecker product, we define
$\tilde{\boldsymbol{X}} = \boldsymbol{Z} \otimes \boldsymbol{1}^T \odot \boldsymbol{X}
\in \mathbb{R}^{n \times p}$ as the design matrix $\boldsymbol{X}$ with
zero-valued rows where observations are missing in $\boldsymbol{Y}$ (where
$\boldsymbol{1}$ represents a vector of ones of dimension $p$).

We estimate the parameters $\boldsymbol{\beta}$ with the least-squares estimator:

\begin{equation}
    \hat{\boldsymbol{\beta}} = \left(\tilde{\boldsymbol{X}}^T \tilde{\boldsymbol{X}}\right)^{-1} \tilde{\boldsymbol{X}}^T \tilde{\boldsymbol{Y}}.
\end{equation}

We compute the estimated residuals as $\hat{\boldsymbol{\varepsilon}} = {\tilde{\boldsymbol{Y}}} - \tilde{{\boldsymbol{X}}} \hat{\boldsymbol{\beta}}$.

We estimate the parameters $\boldsymbol{\vartheta}$ of the missingness process
$\boldsymbol{Z}$ via maximum likelihood, assuming that $\boldsymbol{Z}$ is
generated from a Markov model with the following transition probabilities:

\begin{equation}
    \label{eq:markov_model_def}
    \begin{aligned}
& P\left( Z_2=1 \mid Z_1=1\right)=1 - p_1 \\
& P\left(Z_2=1 \mid Z_1=0\right) = p_2 \\
& P\left(Z_2=0 \mid Z_1=1\right)=p_1 \\
& P\left(Z_2=0 \mid Z_1=0\right)=1-p_2.
\end{aligned}
\end{equation}

We then estimate the parameters $\boldsymbol{\gamma}$ using a Generalized
Method of Wavelet Moments approach [@guerrier2013wavelet].

More precisely, we rely on results from @xu2017study, which provide a
computationally efficient form of the theoretical Allan variance (equivalent
to the Haar wavelet variance up to a constant) for zero-mean stochastic
processes such as $\boldsymbol{\varepsilon}$, avoiding large matrix
multiplications in the objective function. In @xu2017study they generalize the
results in @zhang2008allan to zero-mean non-stationary processes by using
averages of the diagonals and super-diagonals of the covariance matrix of
$\boldsymbol{\varepsilon}$. This implies that the GMWM does not require
storage of the $n \times n$ covariance matrix of $\boldsymbol{\varepsilon}$,
but only a vector of length $n$ that is plugged into an explicit formula
consisting of a linear combination of its elements (these elements are
averages of the diagonal and super-diagonals of the covariance matrix).


# Estimating tectonic velocities and crustal uplift

While the GMWMX, as described above and in more detail in
@voirol2024inference, is a general method for estimating large linear models
with complex dependence structures in the presence of missing data, the
`gmwmx2` R package is currently developed specifically to estimate tectonic
velocities from position time series in graticule distance (GD) coordinates
provided by the Nevada Geodetic Laboratory
[@blewitt2024improved; @blewitt2018harnessing].

## Trajectory model

To estimate the trajectory model (see, e.g., @bevis2014 for more details), we
construct the design matrix $\boldsymbol{X}$ such that the $i$-th component of
the vector $\mathbf{X} {{\boldsymbol{\beta}}}$ can be described as follows, with
$t_i$ representing the $i^{\text{th}}$ ordered time point (epoch) in Modified
Julian Date and $t_0$ representing the reference epoch located exactly in the
middle of the start and end points of the time series:

\begin{split}
 \mathbb{E}[\mathbf{Y}_i] &= \mathbf{X}_i^T {{\boldsymbol{\beta}}}  \\
&= a + b \left(t_{i} - t_0\right) +  \sum_{h=1}^{m}\left[c_{h} \sin \left(2 \pi f_{h} t_{i}\right)  + d_h \cos \left(2 \pi f_h t_i\right)\right] + \\& \sum_{j=1}^{r}e_j H\left(t_i - t_j\right) + \sum_{k = 1 }^{s} l_k \left[1- \exp\left(\frac{-(t_i-t_k)}{\tau_k}\right)\right]H\left(t_{i}-\tau_k\right) 
\end{split}

where $a$ is the initial position at the reference epoch $t_0$, $b$ is the
velocity parameter, and $c_h$ and $d_h$ are the periodic motion parameters
($h = 1$ and $h = 2$ represent the annual and semi-annual seasonal terms,
respectively with $f_1 = \frac{1}{365.25}$ and $f_2 = \frac{2}{365.25}$). The
offset terms model earthquakes, equipment changes, or human intervention, in
which $e_j$ is the magnitude of the step at epochs $t_j$, $r$ is the total
number of offsets, and $H$ is the Heaviside step function defined as
$H(x):= \begin{cases}1, & x \geq 0 \\ 0, & x<0\end{cases}$. The last term allows
us to model post-seismic deformation (see, e.g., @sobrero2020logarithmic) where
$s$ is the number of post-seismic relaxation times specified, $t_k$ is the time
when the relaxation $k$
starts in Modified Julian Date (MJD), $\tau_k$ is the relaxation period in days
for the post-seismic relaxation $k$, and $l_k$ is the amplitude of the
transient. Note that by default the estimates of the functional parameters are
provided in units/day. 

When loading data from a specific station using
`gmwmx2::download_station_ngl()`, we extract from the Nevada Geodetic Laboratory
the position time series in GD coordinates, the time steps associated with
equipment or software changes, and the time steps associated with earthquakes
near the station. All these objects are stored in an object of class
`gnss_ts_ngl`.  

When applying the function `gmwmx2::gmwmx2()` to an object of class
`gnss_ts_ngl`, we construct the design matrix $\boldsymbol{X}$ by considering
an offset term for all equipment or software change steps and all earthquakes
indicated by the NGL. We also specify a post-seismic relaxation term for all
earthquakes indicated by the NGL. If no relaxation time is specified in the
argument `vec_earthquakes_relaxation_time`, we use a default relaxation time of
$365.25$ days. 

## Stochastic model

It is generally recognized that noise in GNSS time series is best described by
a combination of colored noise plus white noise
[@he2017review; @langbein2008noise; @williams2004error; @bos2013fast], where the
white noise generally models noise at high frequencies and the colored noise
models the lower frequencies of the spectrum. 

# Example of estimation
Let us showcase how to estimate tectonic velocity for one specific component
(North, East, or Vertical) of one station.

First, load the `gmwmx2` package.
```{r}
library(gmwmx2)
```

# Download a station from Nevada Geodetic Laboratory
```{r}
station_data <- download_station_ngl("CHML")
```

# Plot Station
```{r}
plot(station_data)
```


# Plot Northing component
```{r}
plot(station_data, component = "N")
```


# Estimate model on station data
```{r}
fit1 <- gmwmx2_new(station_data, n_seasonal = 2, component = "N", model = wn()+ pl())
fit1
```


# Plot estimated model
```{r}
plot(fit1)
```
```{r}
fit2 <- gmwmx2_new(station_data, n_seasonal = 2, component = "N", model = wn()+ flicker())
fit2
```

```{r}
plot(fit2)
```

```{r}
fit3 = gmwmx2_new(station_data, n_seasonal = 2, component = "N", model = wn() + ar1())
fit3
plot(fit3)
```




# References
