---
title: "GMWMX: Estimate linear models with dependent errors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GMWMX: Estimate linear models with dependent errors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 10,
  fig.height = 6,
  fig.align = "center"
)

# so not to executate this vignette during pckg devlopment
# knitr::opts_chunk$set(eval = FALSE)
```

This vignette is a detailed, self-contained simulation walkthrough that showcases
`gmwmx2_new()` for an arbitrary design matrix `X` and response vector `y`. We build the
signal, generate noise, fit models, and run Monte Carlo experiments to check coverage of
confidence intervals. For each setting, we also plot the empirical distributions of the
estimated functional and stochastic parameters.

```{r, message=FALSE,warning=FALSE}
library(gmwmx2)
library(wv)
library(dplyr)

boxplot_mean_dot <- function(x, ...) {
  boxplot(x, ...)
  x_mat <- as.matrix(x)
  mean_vals <- colMeans(x_mat, na.rm = TRUE)
  points(seq_along(mean_vals), mean_vals, pch = 16, col = "black")
}
```

# Overview of the simulation

We consider the linear model
\[
  \mathbf{Y} = \mathbf{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon},
\]
where \(\boldsymbol{\varepsilon} \sim F\!\left(0, \boldsymbol{\Sigma}(\boldsymbol{\gamma})\right)\)
denotes a dependent‑error process with distribution family \(F\) and variance
structure \(\boldsymbol{\Sigma}(\boldsymbol{\gamma})\) parametrized by \(\boldsymbol{\gamma}\).

The goal is to:

1. Fit `gmwmx2_new()` on simulated data with known parameters.
2. Compare results across different stochastic models.
3. Run a Monte Carlo experiment to assess empirical coverage of the confidence
intervals on `beta`.

# Build an arbitrary design matrix X

We use a generic design matrix that mirrors many geodetic or environmental signals:

1. Intercept.
2. Linear trend.
3. Annual sinusoid and cosine.

This is arbitrary and can be replaced with any design matrix suitable to your
application.

```{r}
n = 10*365 
X = matrix(NA, nrow = n, ncol = 4)
# intercept
X[, 1] = 1
# trend
X[, 2] = 1:n
# annual sinusoid
omega_1 <- (1 / 365.25) * 2 * pi
X[, 3] <- sin((1:n) * omega_1)
X[, 4] <- cos((1:n) * omega_1)
beta = c(1, 0.01, 2, 1.5)

# visualize the deterministic signal
plot(x = X[, 2], y = X %*% beta, type = "l",
     main = "Signal", xlab = "Time", ylab = "Signal")
```

We now define three different stochastic noise settings using the same `X` and `beta`.

# Example 1: White noise + AR(1)

This is a classic short-memory model often used as a baseline.

```{r}
phi_ar1 = 0.95
sigma2_ar1 = 20
sigma2_wn = 20
eps = gmwmx2::generate(ar1(phi = phi_ar1, sigma2 = sigma2_ar1) + wn(sigma2_wn), n = n, seed = 123)$series
plot(wv::wvar(eps))
y = X %*% beta + eps
plot(X[, 2], y, type = "l", main = "Simulated y (WN + AR1)")
```

# Fit model (WN + AR1)

We fit using `gmwmx2_new()` with a matching composite model.

```{r}
fit = gmwmx2_new(X = X, y = y, model = wn() + ar1() )
print(fit)
```

To assess uncertainty, we run a Monte Carlo simulation and check empirical coverage
of the confidence intervals for `beta`.

```{r}
B = 100
mat_res = matrix(NA, nrow=B, ncol=19)
for(b in seq(B)){
  eps = generate(ar1(phi=0.95, sigma2=20) + wn(20), n=n, seed = (123 + b))$series
  y = X %*% beta + eps
  fit = gmwmx2_new(X = X, y = y, model = wn() + ar1() )
  # mispecified model assuming white noise as the stochastic model
  fit2 = lm(y~X[,2] + X[,3] + X[,4])

  mat_res[b, ] = c(fit$beta_hat, fit$std_beta_hat,
                   summary(fit2)$coefficients[,1],
                   summary(fit2)$coefficients[,2],
                   fit$theta_domain$`AR(1)_2`,
                   fit$theta_domain$`White Noise_1`)
  # cat("Iteration ", b, " completed.\n")
}
```

```{r}
# compute empirical coverage
mat_res_df = as.data.frame(mat_res)
colnames(mat_res_df) = c("gmwmx_beta0_hat", "gmwmx_beta1_hat",
                         "gmwmx_beta2_hat", "gmwmx_beta3_hat",
                         "gmwmx_std_beta0_hat", "gmwmx_std_beta1_hat",
                         "gmwmx_std_beta2_hat", "gmwmx_std_beta3_hat",
                         "lm_beta0_hat", "lm_beta1_hat", "lm_beta2_hat", "lm_beta3_hat",
                         "lm_std_beta0_hat", "lm_std_beta1_hat", "lm_std_beta2_hat", "lm_std_beta3_hat",
                         "phi_ar1","sigma_2_ar1" ,"sigma_2_wn")

```

# Plot empirical distributions (beta and stochastic parameters)

```{r}
par(mfrow = c(1, 4))
boxplot_mean_dot(mat_res_df[, c("gmwmx_beta0_hat")],las=1,
        names = c("beta0"),
        main = expression(beta[0]), ylab = "Estimate")
abline(h = beta[1], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta1_hat")],las=1,
        names = c("beta1"),
        main = expression(beta[1]), ylab = "Estimate")
abline(h = beta[2], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta2_hat")],las=1,
        names = c("beta2"),
        main = expression(beta[2]), ylab = "Estimate")
abline(h = beta[3], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta3_hat")],las=1,
        names = c("beta3"),
        main = expression(beta[3]), ylab = "Estimate")
abline(h = beta[4], col = "black", lwd = 2)


par(mfrow = c(1, 3))

boxplot_mean_dot(mat_res_df$phi_ar1,las=1,
        names = c("phi_ar1"),
        main = expression(phi["AR1"]), ylab = "Estimate")
abline(h = phi_ar1, col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df$sigma_2_ar1,las=1,
        names = c("phi_ar1"),
        main = expression(sigma["AR1"]^2), ylab = "Estimate")
abline(h = sigma2_ar1, col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df$sigma_2_wn,las=1,
        names = c("phi_ar1"),
        main = expression(sigma["WN"]^2), ylab = "Estimate")
abline(h = sigma2_wn, col = "black", lwd = 2)
par(mfrow = c(1, 1))
```

# Compute empirical coverage of confidence intervals for beta
```{r}
zval = qnorm(0.975)
mat_res_df$upper_ci_gmwmx_beta0 = mat_res_df$gmwmx_beta0_hat + zval * mat_res_df$gmwmx_std_beta0_hat
mat_res_df$lower_ci_gmwmx_beta0 = mat_res_df$gmwmx_beta0_hat - zval * mat_res_df$gmwmx_std_beta0_hat
mat_res_df$upper_ci_gmwmx_beta1 = mat_res_df$gmwmx_beta1_hat + zval * mat_res_df$gmwmx_std_beta1_hat
mat_res_df$lower_ci_gmwmx_beta1 = mat_res_df$gmwmx_beta1_hat - zval * mat_res_df$gmwmx_std_beta1_hat
# empirical coverage of gmwmx beta
dplyr::between(rep(beta[1], B), mat_res_df$lower_ci_gmwmx_beta0, mat_res_df$upper_ci_gmwmx_beta0) %>% mean()
dplyr::between(rep(beta[2], B), mat_res_df$lower_ci_gmwmx_beta1, mat_res_df$upper_ci_gmwmx_beta1) %>% mean()

# do the same for lm beta
mat_res_df$upper_ci_lm_beta0 = mat_res_df$lm_beta0_hat + zval * mat_res_df$lm_std_beta0_hat
mat_res_df$lower_ci_lm_beta0 = mat_res_df$lm_beta0_hat - zval * mat_res_df$lm_std_beta0_hat
mat_res_df$upper_ci_lm_beta1 = mat_res_df$lm_beta1_hat + zval * mat_res_df$lm_std_beta1_hat
mat_res_df$lower_ci_lm_beta1 = mat_res_df$lm_beta1_hat - zval * mat_res_df$lm_std_beta1_hat
dplyr::between(rep(beta[1], B), mat_res_df$lower_ci_lm_beta0, mat_res_df$upper_ci_lm_beta0) %>% mean()
dplyr::between(rep(beta[2], B), mat_res_df$lower_ci_lm_beta1, mat_res_df$upper_ci_lm_beta1) %>% mean()
```



# Example 2: White noise + stationary power-law

Here we replace AR(1) by a stationary power-law component with spectral index
`kappa = -0.8`. This introduces long-memory behavior.

```{r}
kappa_pl <- -0.8
sigma2_wn <- 10
sigma2_pl <- 10
eps_pl = generate(wn(sigma2_wn) + pl(kappa = kappa_pl, sigma2 = sigma2_pl),
                  n = n, seed = 123)
plot(eps_pl$series, type = "l")
plot(wv::wvar(eps_pl$series))
y_pl = X %*% beta + eps_pl$series

fit_pl = gmwmx2_new(X = X, y = y_pl, model = wn() + pl())
fit_pl
```

Monte Carlo coverage for the power-law case:

```{r}
B_pl = 100
mat_res_pl = matrix(NA, nrow = B_pl, ncol = 19)
for(b in seq(B_pl)){
  eps = generate(wn(sigma2_wn) + pl(kappa = kappa_pl, sigma2 = sigma2_pl),
                 n = n, seed = (123 + b))$series
  y = X %*% beta + eps
  fit = gmwmx2_new(X = X, y = y, model = wn() + pl())
  fit2 = lm(y~X[,2] + X[,3] + X[,4])

  mat_res_pl[b, ] = c(fit$beta_hat, fit$std_beta_hat,
                      summary(fit2)$coefficients[,1],
                      summary(fit2)$coefficients[,2],
                      fit$theta_domain$`Stationary PowerLaw_2`,
                      fit$theta_domain$`White Noise_1`)
  # cat("Iteration ", b, " \n")
}

mat_res_pl_df = as.data.frame(mat_res_pl)
colnames(mat_res_pl_df) = c("gmwmx_beta0_hat", "gmwmx_beta1_hat",
                            "gmwmx_beta2_hat", "gmwmx_beta3_hat",
                            "gmwmx_std_beta0_hat", "gmwmx_std_beta1_hat",
                            "gmwmx_std_beta2_hat", "gmwmx_std_beta3_hat",
                            "lm_beta0_hat", "lm_beta1_hat", "lm_beta2_hat", "lm_beta3_hat",
                            "lm_std_beta0_hat", "lm_std_beta1_hat",
                            "lm_std_beta2_hat", "lm_std_beta3_hat",
                            "kappa_pl","sigma2_pl" ,"sigma2_wn")
```

# Plot empirical distributions (beta and stochastic parameters)

```{r}
par(mfrow = c(1, 4))
boxplot_mean_dot(mat_res_df[, c("gmwmx_beta0_hat")],las=1,
        names = c("beta0"),
        main = expression(beta[0]), ylab = "Estimate")
abline(h = beta[1], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta1_hat")],las=1,
        names = c("beta1"),
        main = expression(beta[1]), ylab = "Estimate")
abline(h = beta[2], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta2_hat")],las=1,
        names = c("beta2"),
        main = expression(beta[2]), ylab = "Estimate")
abline(h = beta[3], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta3_hat")],las=1,
        names = c("beta3"),
        main = expression(beta[3]), ylab = "Estimate")
abline(h = beta[4], col = "black", lwd = 2)

par(mfrow = c(1, 1))
```


```{r}
par(mfrow = c(1, 3))

boxplot_mean_dot(mat_res_pl_df$kappa_pl,las=1,
        main = expression(kappa["PL"]), ylab = "Estimate")
abline(h = kappa_pl, col = "black", lwd = 2)

boxplot_mean_dot(mat_res_pl_df$sigma2_pl,las=1,
        main = expression(sigma["PL"]^2), ylab = "Estimate")
abline(h = sigma2_pl, col = "black", lwd = 2)

boxplot_mean_dot(mat_res_pl_df$sigma2_wn,las=1,
        names = c("phi_ar1"),
        main = expression(sigma["WN"]^2), ylab = "Estimate")
abline(h = sigma2_wn, col = "black", lwd = 2)
par(mfrow = c(1, 1))
```


# Compute empirical coverage of confidence intervals for beta
```{r}
zval = qnorm(0.975)
mat_res_pl_df$upper_ci_gmwmx_beta0 = mat_res_pl_df$gmwmx_beta0_hat + zval * mat_res_pl_df$gmwmx_std_beta0_hat
mat_res_pl_df$lower_ci_gmwmx_beta0 = mat_res_pl_df$gmwmx_beta0_hat - zval * mat_res_pl_df$gmwmx_std_beta0_hat
mat_res_pl_df$upper_ci_gmwmx_beta1 = mat_res_pl_df$gmwmx_beta1_hat + zval * mat_res_pl_df$gmwmx_std_beta1_hat
mat_res_pl_df$lower_ci_gmwmx_beta1 = mat_res_pl_df$gmwmx_beta1_hat - zval * mat_res_pl_df$gmwmx_std_beta1_hat
# empirical coverage of gmwmx beta
dplyr::between(rep(beta[1], B_pl), mat_res_pl_df$lower_ci_gmwmx_beta0, mat_res_pl_df$upper_ci_gmwmx_beta0) %>% mean()
dplyr::between(rep(beta[2], B_pl), mat_res_pl_df$lower_ci_gmwmx_beta1, mat_res_pl_df$upper_ci_gmwmx_beta1) %>% mean()

# do the same for lm beta
mat_res_pl_df$upper_ci_lm_beta0 = mat_res_pl_df$lm_beta0_hat + zval * mat_res_pl_df$lm_std_beta0_hat
mat_res_pl_df$lower_ci_lm_beta0 = mat_res_pl_df$lm_beta0_hat - zval * mat_res_pl_df$lm_std_beta0_hat
mat_res_pl_df$upper_ci_lm_beta1 = mat_res_pl_df$lm_beta1_hat + zval * mat_res_pl_df$lm_std_beta1_hat
mat_res_pl_df$lower_ci_lm_beta1 = mat_res_pl_df$lm_beta1_hat - zval * mat_res_pl_df$lm_std_beta1_hat
dplyr::between(rep(beta[1], B_pl), mat_res_pl_df$lower_ci_lm_beta0, mat_res_pl_df$upper_ci_lm_beta0) %>% mean()
dplyr::between(rep(beta[2], B_pl), mat_res_pl_df$lower_ci_lm_beta1, mat_res_pl_df$upper_ci_lm_beta1) %>% mean()
```

# Example 3: White noise + flicker

Flicker noise corresponds to a non stationary power-law with spectral index `kappa = -1` but is
handled explicitly by `flicker()` in this package.

```{r}
sigma2_wn_fl <- 15
sigma2_fl <- 10
eps_fl = generate(wn(sigma2_wn_fl) + flicker(sigma2 = sigma2_fl),
                  n = n, seed = 123)
plot(eps_fl$series, type = "l")
plot(wv::wvar(eps_fl$series))
y_fl = X %*% beta + eps_fl$series

fit_fl = gmwmx2_new(X = X, y = y_fl, model = wn() + flicker())
fit_fl
```

Monte Carlo coverage for the flicker case:

```{r}
B_fl = 100
mat_res_fl = matrix(NA, nrow = B_fl, ncol = 18)
for(b in seq(B_fl)){
  eps = generate(wn(sigma2_wn_fl) + flicker(sigma2 = sigma2_fl),
                 n = n, seed = (123 + b))$series
  y = X %*% beta + eps
  fit = gmwmx2_new(X = X, y = y, model = wn() + flicker())
  fit2 = lm(y~X[,2] + X[,3] + X[,4])

  mat_res_fl[b, ] = c(fit$beta_hat, fit$std_beta_hat,
                      summary(fit2)$coefficients[,1],
                      summary(fit2)$coefficients[,2],
                      fit$theta_domain$`Flicker_2`,
                      fit$theta_domain$`White Noise_1`)
  # cat("Iteration ", b, " \n")
}

mat_res_fl_df = as.data.frame(mat_res_fl)
colnames(mat_res_fl_df) = c("gmwmx_beta0_hat", "gmwmx_beta1_hat",
                            "gmwmx_beta2_hat", "gmwmx_beta3_hat",
                            "gmwmx_std_beta0_hat", "gmwmx_std_beta1_hat",
                            "gmwmx_std_beta2_hat", "gmwmx_std_beta3_hat",
                            "lm_beta0_hat", "lm_beta1_hat", "lm_beta2_hat", "lm_beta3_hat",
                            "lm_std_beta0_hat", "lm_std_beta1_hat",
                            "lm_std_beta2_hat", "lm_std_beta3_hat",
                            "sigma2_fl" ,"sigma2_wn")
```


# Compute empirical coverage of confidence intervals for beta
```{r}
zval = qnorm(0.975)
mat_res_fl_df$upper_ci_gmwmx_beta0 = mat_res_fl_df$gmwmx_beta0_hat + zval * mat_res_fl_df$gmwmx_std_beta0_hat
mat_res_fl_df$lower_ci_gmwmx_beta0 = mat_res_fl_df$gmwmx_beta0_hat - zval * mat_res_fl_df$gmwmx_std_beta0_hat
mat_res_fl_df$upper_ci_gmwmx_beta1 = mat_res_fl_df$gmwmx_beta1_hat + zval * mat_res_fl_df$gmwmx_std_beta1_hat
mat_res_fl_df$lower_ci_gmwmx_beta1 = mat_res_fl_df$gmwmx_beta1_hat - zval * mat_res_fl_df$gmwmx_std_beta1_hat
# empirical coverage of gmwmx beta
dplyr::between(rep(beta[1], B_pl), mat_res_fl_df$lower_ci_gmwmx_beta0, mat_res_fl_df$upper_ci_gmwmx_beta0) %>% mean()
dplyr::between(rep(beta[2], B_pl), mat_res_fl_df$lower_ci_gmwmx_beta1, mat_res_fl_df$upper_ci_gmwmx_beta1) %>% mean()

# do the same for lm beta
mat_res_fl_df$upper_ci_lm_beta0 = mat_res_fl_df$lm_beta0_hat + zval * mat_res_fl_df$lm_std_beta0_hat
mat_res_fl_df$lower_ci_lm_beta0 = mat_res_fl_df$lm_beta0_hat - zval * mat_res_fl_df$lm_std_beta0_hat
mat_res_fl_df$upper_ci_lm_beta1 = mat_res_fl_df$lm_beta1_hat + zval * mat_res_fl_df$lm_std_beta1_hat
mat_res_fl_df$lower_ci_lm_beta1 = mat_res_fl_df$lm_beta1_hat - zval * mat_res_fl_df$lm_std_beta1_hat
dplyr::between(rep(beta[1], B_pl), mat_res_fl_df$lower_ci_lm_beta0, mat_res_fl_df$upper_ci_lm_beta0) %>% mean()
dplyr::between(rep(beta[2], B_pl), mat_res_fl_df$lower_ci_lm_beta1, mat_res_fl_df$upper_ci_lm_beta1) %>% mean()
```



# Plot empirical distributions (beta and stochastic parameters)

```{r}
par(mfrow = c(1, 4))
boxplot_mean_dot(mat_res_df[, c("gmwmx_beta0_hat")],las=1,
        names = c("beta0"),
        main = expression(beta[0]), ylab = "Estimate")
abline(h = beta[1], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta1_hat")],las=1,
        names = c("beta1"),
        main = expression(beta[1]), ylab = "Estimate")
abline(h = beta[2], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta2_hat")],las=1,
        names = c("beta2"),
        main = expression(beta[2]), ylab = "Estimate")
abline(h = beta[3], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta3_hat")],las=1,
        names = c("beta3"),
        main = expression(beta[3]), ylab = "Estimate")
abline(h = beta[4], col = "black", lwd = 2)

par(mfrow = c(1, 1))
```


```{r}
par(mfrow = c(1, 2))

boxplot_mean_dot(mat_res_fl_df$sigma2_fl,las=1,
        main = expression(sigma["FL"]^2), ylab = "Estimate")
abline(h = sigma2_fl, col = "black", lwd = 2)

boxplot_mean_dot(mat_res_fl_df$sigma2_wn,las=1,
        names = c("phi_ar1"),
        main = expression(sigma["WN"]^2), ylab = "Estimate")
abline(h = sigma2_wn_fl, col = "black", lwd = 2)
par(mfrow = c(1, 1))
```





# Interpreting the outputs

The `gmwmx2_new()` object contains:

1. `beta_hat` and `std_beta_hat`: regression estimates and standard errors.
2. `theta_domain`: estimated stochastic parameters (e.g., AR(1), WN variance, power-law
parameters, flicker variance).
3. Diagnostic information about the fitted wavelet variance and optimization output.

The Monte Carlo results provide empirical coverage for the nominal 95% confidence
intervals. This helps assess whether standard errors from `gmwmx2_new()` are well
calibrated under different noise models.


