---
title: "GMWMX: Estimate linear models with dependent errors and missing observations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{GMWMX: Estimate linear models with dependent errors and missing observations}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette shows how to perform inference for a linear regression model
with dependent errors and missing observations. We consider the model
\[
  \mathbf{Y} = \mathbf{X}\boldsymbol{\beta} + \boldsymbol{\varepsilon},
\]
where the error process \(\boldsymbol{\varepsilon}\) follows a composite stochastic model. Missingness is modeled by a binary process \(\mathbf{Z} \in \{0,1\}^n\) with \(Z_t = 1\) indicating
observation, and \(Z_t = 0\) a missing observation. The observed series is
\[
  \tilde{\mathbf{Y}} = \mathbf{Y} \circ \mathbf{Z},
\]
which we implement by setting \(\tilde{\mathbf{Y}}\) to `NA` when \(Z_t = 0\).

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 10,
  fig.height = 6,
  fig.align = "center"
)

# so not to executate this vignette during pckg devlopment
# knitr::opts_chunk$set(eval = FALSE)
```

Overall, these examples illustrate how `gmwmx2_new()` supports inference for
linear regression models with dependent errors when observations are missing,
and how the results compare to a naive OLS fit that ignores dependence.


```{r}
library(gmwmx2)
library(wv)
library(dplyr)

boxplot_mean_dot <- function(x, ...) {
  boxplot(x, ...)
  x_mat <- as.matrix(x)
  mean_vals <- colMeans(x_mat, na.rm = TRUE)
  points(seq_along(mean_vals), mean_vals, pch = 16, col = "black")
}
```

We load the packages used for simulation, wavelet variance diagnostics, and
summary plots. The helper `boxplot_mean_dot()` overlays Monte Carlo means on
top of each boxplot for easier comparison.

We first construct a design matrix with an intercept, linear trend, and annual
seasonal components. This deterministic signal will be used across examples.

```{r}
n = 10*365
X = matrix(NA, nrow = n, ncol = 4)
# intercept
X[, 1] = 1
# trend
X[, 2] = 1:n
# annual sinusoid
omega_1 <- (1 / 365.25) * 2 * pi
X[, 3] <- sin((1:n) * omega_1)
X[, 4] <- cos((1:n) * omega_1)
beta = c(1, 0.01, 2, 1.5)

# visualize the deterministic signal
plot(x = X[, 2], y = X %*% beta, type = "l",
     main = "Signal", xlab = "Time", ylab = "Signal")
```

Next we simulate dependent errors, add them to the signal, and then introduce
missing observations using a two‑state Markov model. This is the key difference
from `vignettes/gmwmx2_new.Rmd`, which uses fully observed data.

```{r}
# Example 1: White noise + AR(1)


phi_ar1 = 0.95
sigma2_ar1 = 20
sigma2_wn = 20
eps = generate(ar1(phi = phi_ar1, sigma2 = sigma2_ar1) + wn(sigma2_wn), n = n, seed = 123)$series
plot(wv::wvar(eps))
y = X %*% beta + eps
# generate missingness
mod_missing = markov_two_states(p1 = 0.05, p2 = .95)
Z = generate(mod_missing, n = n, seed = 123)
plot(Z)
Z_process = Z$series
y_observed = ifelse(Z_process == 1, y, NA)
plot(X[, 2], y_observed, type = "l", main = "Simulated y (WN + AR1)")
# add grey boxes at missing indices (x ± eps) to visualize NA locations;
# unlike `vignettes/gmwmx2_new.Rmd`, which plots the fully observed series
na_idx <- which(is.na(y_observed))

# choose epsilon relative to x spacing
dx <- median(diff(X[, 2]), na.rm = TRUE)
eps <- 0.45 * dx

ylim <- par("usr")[3:4]
for (i in na_idx) {
  rect(
    xleft = X[i, 2] - eps,
    xright = X[i, 2] + eps,
    ybottom = ylim[1],
    ytop = ylim[2],
    col = adjustcolor("grey20", alpha.f = 0.5),
    border = NA
  )
}

lines(X[, 2], y_observed)
```

We then fit the dependent‑error regression using `gmwmx2_new()`, which accounts
for the stochastic error model and the missing observations.



```{r}
#
# Fit model (WN + AR1)

fit = gmwmx2_new(X = X, y = y_observed, model = wn() + ar1() )
print(fit)




B = 100
mat_res = matrix(NA, nrow=B, ncol=19)
for(b in seq(B)){

  eps = generate(ar1(phi=0.95, sigma2=20) + wn(20), n=n, seed = (123 + b))$series
  y = X %*% beta + eps
  Z = generate(mod_missing, n = n, seed = 123 + b)$series
  y_observed = ifelse(Z == 1, y, NA)
  fit = gmwmx2_new(X = X, y = y_observed, model = wn() + ar1() )
  # mispecified model assuming white noise as the stochastic model and only on observed data
  fit2 = lm(y_observed~X[,2] + X[,3] + X[,4])

  mat_res[b, ] = c(fit$beta_hat, fit$std_beta_hat,
                   summary(fit2)$coefficients[,1],
                   summary(fit2)$coefficients[,2],
                   fit$theta_domain$`AR(1)_2`,
                   fit$theta_domain$`White Noise_1`)
  # cat("Iteration ", b, " completed.\n")
}
```

We now assess inference quality under missingness by repeating the simulation,
fitting the correct dependent‑error model with `gmwmx2_new()`, and comparing
against a misspecified OLS fit that treats the errors as i.i.d. and uses only
observed data.

```{r}
# compute empirical coverage
mat_res_df = as.data.frame(mat_res)
colnames(mat_res_df) = c("gmwmx_beta0_hat", "gmwmx_beta1_hat",
                         "gmwmx_beta2_hat", "gmwmx_beta3_hat",
                         "gmwmx_std_beta0_hat", "gmwmx_std_beta1_hat",
                         "gmwmx_std_beta2_hat", "gmwmx_std_beta3_hat",
                         "lm_beta0_hat", "lm_beta1_hat", "lm_beta2_hat", "lm_beta3_hat",
                         "lm_std_beta0_hat", "lm_std_beta1_hat", "lm_std_beta2_hat", "lm_std_beta3_hat",
                         "phi_ar1","sigma_2_ar1" ,"sigma_2_wn")



# Plot empirical distributions (beta and stochastic parameters)

par(mfrow = c(1, 4))
boxplot_mean_dot(mat_res_df[, c("gmwmx_beta0_hat")],las=1,
                 names = c("beta0"),
                 main = expression(beta[0]), ylab = "Estimate")
abline(h = beta[1], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta1_hat")],las=1,
                 names = c("beta1"),
                 main = expression(beta[1]), ylab = "Estimate")
abline(h = beta[2], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta2_hat")],las=1,
                 names = c("beta2"),
                 main = expression(beta[2]), ylab = "Estimate")
abline(h = beta[3], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta3_hat")],las=1,
                 names = c("beta3"),
                 main = expression(beta[3]), ylab = "Estimate")
abline(h = beta[4], col = "black", lwd = 2)


par(mfrow = c(1, 3))

boxplot_mean_dot(mat_res_df$phi_ar1,las=1,
                 names = c("phi_ar1"),
                 main = expression(phi["AR1"]), ylab = "Estimate")
abline(h = phi_ar1, col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df$sigma_2_ar1,las=1,
                 names = c("phi_ar1"),
                 main = expression(sigma["AR1"]^2), ylab = "Estimate")
abline(h = sigma2_ar1, col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df$sigma_2_wn,las=1,
                 names = c("phi_ar1"),
                 main = expression(sigma["WN"]^2), ylab = "Estimate")
abline(h = sigma2_wn, col = "black", lwd = 2)
par(mfrow = c(1, 1))
```

The boxplots summarize the empirical distribution of the regression estimates
and stochastic parameters across Monte Carlo replications, with the true values
shown as horizontal lines.

Coverage is computed using normal approximations for the coefficient intervals.
We report the empirical fraction of times the true \(\boldsymbol{\beta}\)
entries fall inside the estimated intervals for both `gmwmx2_new()` and OLS.


```{r}
# Compute empirical coverage of confidence intervals for beta

zval = qnorm(0.975)
mat_res_df$upper_ci_gmwmx_beta0 = mat_res_df$gmwmx_beta0_hat + zval * mat_res_df$gmwmx_std_beta0_hat
mat_res_df$lower_ci_gmwmx_beta0 = mat_res_df$gmwmx_beta0_hat - zval * mat_res_df$gmwmx_std_beta0_hat
mat_res_df$upper_ci_gmwmx_beta1 = mat_res_df$gmwmx_beta1_hat + zval * mat_res_df$gmwmx_std_beta1_hat
mat_res_df$lower_ci_gmwmx_beta1 = mat_res_df$gmwmx_beta1_hat - zval * mat_res_df$gmwmx_std_beta1_hat
# empirical coverage of gmwmx beta
dplyr::between(rep(beta[1], B), mat_res_df$lower_ci_gmwmx_beta0, mat_res_df$upper_ci_gmwmx_beta0) %>% mean()
dplyr::between(rep(beta[2], B), mat_res_df$lower_ci_gmwmx_beta1, mat_res_df$upper_ci_gmwmx_beta1) %>% mean()

# do the same for lm beta
mat_res_df$upper_ci_lm_beta0 = mat_res_df$lm_beta0_hat + zval * mat_res_df$lm_std_beta0_hat
mat_res_df$lower_ci_lm_beta0 = mat_res_df$lm_beta0_hat - zval * mat_res_df$lm_std_beta0_hat
mat_res_df$upper_ci_lm_beta1 = mat_res_df$lm_beta1_hat + zval * mat_res_df$lm_std_beta1_hat
mat_res_df$lower_ci_lm_beta1 = mat_res_df$lm_beta1_hat - zval * mat_res_df$lm_std_beta1_hat
dplyr::between(rep(beta[1], B), mat_res_df$lower_ci_lm_beta0, mat_res_df$upper_ci_lm_beta0) %>% mean()
dplyr::between(rep(beta[2], B), mat_res_df$lower_ci_lm_beta1, mat_res_df$upper_ci_lm_beta1) %>% mean()
```

We now repeat the workflow for a long‑memory error model (white noise +
stationary power‑law). The steps mirror the AR(1) case: simulate with
missingness, fit `gmwmx2_new()`, and compare to a misspecified OLS fit.

The parameter \(\kappa\) controls the long‑memory behavior of the power‑law
component, while the variance terms split energy between white noise and the
power‑law process.


```{r}
# Example 2: White noise + stationary power-law



kappa_pl <- -0.8
sigma2_wn <- 15
sigma2_pl <- 5
eps_pl = generate(wn(sigma2_wn) + pl(kappa = kappa_pl, sigma2 = sigma2_pl),
                  n = n, seed = 1234)
plot(eps_pl$series, type = "l")
plot(wv::wvar(eps_pl$series))
y_pl = X %*% beta + eps_pl$series
Z = generate(mod_missing, n = n, seed = 1234)
plot(Z)
Z_process = Z$series
y_observed = ifelse(Z_process == 1, y_pl, NA)


fit_pl = gmwmx2_new(X = X, y = y_observed, model = wn() + pl())
fit_pl
```

As before, we keep the same missingness mechanism so that differences are driven
by the error model rather than by the observation pattern.

Monte Carlo replication for the power‑law case.


```{r}
B_pl = 100
mat_res_pl = matrix(NA, nrow = B_pl, ncol = 19)
for(b in seq(B_pl)){
  eps = generate(wn(sigma2_wn) + pl(kappa = kappa_pl, sigma2 = sigma2_pl),
                 n = n, seed = (123 + b))$series
  y = X %*% beta + eps

  Z = generate(mod_missing, n = n, seed = (123 + b))
  Z_process = Z$series
  y_observed = ifelse(Z_process == 1, y, NA)

  fit = gmwmx2_new(X = X, y = y_observed, model = wn() + pl())
  fit2 = lm(y_observed~X[,2] + X[,3] + X[,4])

  mat_res_pl[b, ] = c(fit$beta_hat, fit$std_beta_hat,
                      summary(fit2)$coefficients[,1],
                      summary(fit2)$coefficients[,2],
                      fit$theta_domain$`Stationary PowerLaw_2`,
                      fit$theta_domain$`White Noise_1`)
  # cat("Iteration ", b, " \n")
}
```

Aggregate the Monte Carlo results and visualize regression estimates.

The next plots focus on the distribution of \(\boldsymbol{\beta}\) estimates
under the power‑law error model.


```{r}
mat_res_pl_df = as.data.frame(mat_res_pl)
colnames(mat_res_pl_df) = c("gmwmx_beta0_hat", "gmwmx_beta1_hat",
                            "gmwmx_beta2_hat", "gmwmx_beta3_hat",
                            "gmwmx_std_beta0_hat", "gmwmx_std_beta1_hat",
                            "gmwmx_std_beta2_hat", "gmwmx_std_beta3_hat",
                            "lm_beta0_hat", "lm_beta1_hat", "lm_beta2_hat", "lm_beta3_hat",
                            "lm_std_beta0_hat", "lm_std_beta1_hat",
                            "lm_std_beta2_hat", "lm_std_beta3_hat",
                            "kappa_pl","sigma2_pl" ,"sigma2_wn")


# Plot empirical distributions (beta and stochastic parameters)


par(mfrow = c(1, 4))
boxplot_mean_dot(mat_res_df[, c("gmwmx_beta0_hat")],las=1,
                 names = c("beta0"),
                 main = expression(beta[0]), ylab = "Estimate")
abline(h = beta[1], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta1_hat")],las=1,
                 names = c("beta1"),
                 main = expression(beta[1]), ylab = "Estimate")
abline(h = beta[2], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta2_hat")],las=1,
                 names = c("beta2"),
                 main = expression(beta[2]), ylab = "Estimate")
abline(h = beta[3], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta3_hat")],las=1,
                 names = c("beta3"),
                 main = expression(beta[3]), ylab = "Estimate")
abline(h = beta[4], col = "black", lwd = 2)

par(mfrow = c(1, 1))
```

Plot the stochastic parameter estimates for the power‑law model.

These plots focus on the long‑memory parameter and variance components that
define the power‑law error process.


```{r}
par(mfrow = c(1, 3))

boxplot_mean_dot(mat_res_pl_df$kappa_pl,las=1,
                 main = expression(kappa["PL"]), ylab = "Estimate")
abline(h = kappa_pl, col = "black", lwd = 2)

boxplot_mean_dot(mat_res_pl_df$sigma2_pl,las=1,
                 main = expression(sigma["PL"]^2), ylab = "Estimate")
abline(h = sigma2_pl, col = "black", lwd = 2)

boxplot_mean_dot(mat_res_pl_df$sigma2_wn,las=1,
                 names = c("phi_ar1"),
                 main = expression(sigma["WN"]^2), ylab = "Estimate")
abline(h = sigma2_wn, col = "black", lwd = 2)
par(mfrow = c(1, 1))
```

Compute empirical coverage for the regression coefficients under the power‑law
error model and compare against OLS.


```{r}
# Compute empirical coverage of confidence intervals for beta

zval = qnorm(0.975)
mat_res_pl_df$upper_ci_gmwmx_beta0 = mat_res_pl_df$gmwmx_beta0_hat + zval * mat_res_pl_df$gmwmx_std_beta0_hat
mat_res_pl_df$lower_ci_gmwmx_beta0 = mat_res_pl_df$gmwmx_beta0_hat - zval * mat_res_pl_df$gmwmx_std_beta0_hat
mat_res_pl_df$upper_ci_gmwmx_beta1 = mat_res_pl_df$gmwmx_beta1_hat + zval * mat_res_pl_df$gmwmx_std_beta1_hat
mat_res_pl_df$lower_ci_gmwmx_beta1 = mat_res_pl_df$gmwmx_beta1_hat - zval * mat_res_pl_df$gmwmx_std_beta1_hat
# empirical coverage of gmwmx beta
dplyr::between(rep(beta[1], B_pl), mat_res_pl_df$lower_ci_gmwmx_beta0, mat_res_pl_df$upper_ci_gmwmx_beta0) %>% mean()
dplyr::between(rep(beta[2], B_pl), mat_res_pl_df$lower_ci_gmwmx_beta1, mat_res_pl_df$upper_ci_gmwmx_beta1) %>% mean()

# do the same for lm beta
mat_res_pl_df$upper_ci_lm_beta0 = mat_res_pl_df$lm_beta0_hat + zval * mat_res_pl_df$lm_std_beta0_hat
mat_res_pl_df$lower_ci_lm_beta0 = mat_res_pl_df$lm_beta0_hat - zval * mat_res_pl_df$lm_std_beta0_hat
mat_res_pl_df$upper_ci_lm_beta1 = mat_res_pl_df$lm_beta1_hat + zval * mat_res_pl_df$lm_std_beta1_hat
mat_res_pl_df$lower_ci_lm_beta1 = mat_res_pl_df$lm_beta1_hat - zval * mat_res_pl_df$lm_std_beta1_hat
dplyr::between(rep(beta[1], B_pl), mat_res_pl_df$lower_ci_lm_beta0, mat_res_pl_df$upper_ci_lm_beta0) %>% mean()
dplyr::between(rep(beta[2], B_pl), mat_res_pl_df$lower_ci_lm_beta1, mat_res_pl_df$upper_ci_lm_beta1) %>% mean()
```

Finally, we repeat the missing‑data inference workflow for a flicker‑noise
error model.

Flicker noise (1/f‑type behavior) provides a contrasting long‑memory structure
to the power‑law case, and helps illustrate how inference changes across
different dependent‑error models.


```{r}
# Example 3: White noise + flicker



sigma2_wn_fl <- 15
sigma2_fl <- 10
eps_fl = generate(wn(sigma2_wn_fl) + flicker(sigma2 = sigma2_fl),
                  n = n, seed = 123)
plot(eps_fl$series, type = "l")
plot(wv::wvar(eps_fl$series))
Z = generate(mod_missing, n = n, seed = 1234)
plot(Z)

y_fl = X %*% beta + eps_fl$series

Z = generate(mod_missing, n = n, seed = 1234)
plot(Z)
Z_process = Z$series


y_observed = ifelse(Z_process == 1, y_fl, NA)

fit_fl = gmwmx2_new(X = X, y = y_observed, model = wn() + flicker())
fit_fl
```

Monte Carlo replication for the flicker‑noise case.


```{r}
B_fl = 100
mat_res_fl = matrix(NA, nrow = B_fl, ncol = 18)
for(b in seq(B_fl)){
  eps = generate(wn(sigma2_wn_fl) + flicker(sigma2 = sigma2_fl),
                 n = n, seed = (123 + b))$series
  y = X %*% beta + eps


  Z = generate(mod_missing, n = n, seed = (123 + b))
  Z_process = Z$series
  y_observed = ifelse(Z_process == 1, y, NA)



  fit = gmwmx2_new(X = X, y = y_observed, model = wn() + flicker())
  fit2 = lm(y_observed~X[,2] + X[,3] + X[,4])

  mat_res_fl[b, ] = c(fit$beta_hat, fit$std_beta_hat,
                      summary(fit2)$coefficients[,1],
                      summary(fit2)$coefficients[,2],
                      fit$theta_domain$`Flicker_2`,
                      fit$theta_domain$`White Noise_1`)
  cat("Iteration ", b, " \n")
}
```


```{r}
mat_res_fl_df = as.data.frame(mat_res_fl)
colnames(mat_res_fl_df) = c("gmwmx_beta0_hat", "gmwmx_beta1_hat",
                            "gmwmx_beta2_hat", "gmwmx_beta3_hat",
                            "gmwmx_std_beta0_hat", "gmwmx_std_beta1_hat",
                            "gmwmx_std_beta2_hat", "gmwmx_std_beta3_hat",
                            "lm_beta0_hat", "lm_beta1_hat", "lm_beta2_hat", "lm_beta3_hat",
                            "lm_std_beta0_hat", "lm_std_beta1_hat",
                            "lm_std_beta2_hat", "lm_std_beta3_hat",
                            "sigma2_fl" ,"sigma2_wn")
```

Compute empirical coverage for the flicker‑noise case.


```{r}
# Compute empirical coverage of confidence intervals for beta

zval = qnorm(0.975)
mat_res_fl_df$upper_ci_gmwmx_beta0 = mat_res_fl_df$gmwmx_beta0_hat + zval * mat_res_fl_df$gmwmx_std_beta0_hat
mat_res_fl_df$lower_ci_gmwmx_beta0 = mat_res_fl_df$gmwmx_beta0_hat - zval * mat_res_fl_df$gmwmx_std_beta0_hat
mat_res_fl_df$upper_ci_gmwmx_beta1 = mat_res_fl_df$gmwmx_beta1_hat + zval * mat_res_fl_df$gmwmx_std_beta1_hat
mat_res_fl_df$lower_ci_gmwmx_beta1 = mat_res_fl_df$gmwmx_beta1_hat - zval * mat_res_fl_df$gmwmx_std_beta1_hat
# empirical coverage of gmwmx beta
dplyr::between(rep(beta[1], B_fl), mat_res_fl_df$lower_ci_gmwmx_beta0, mat_res_fl_df$upper_ci_gmwmx_beta0) %>% mean()
dplyr::between(rep(beta[2], B_fl), mat_res_fl_df$lower_ci_gmwmx_beta1, mat_res_fl_df$upper_ci_gmwmx_beta1) %>% mean()

# do the same for lm beta
mat_res_fl_df$upper_ci_lm_beta0 = mat_res_fl_df$lm_beta0_hat + zval * mat_res_fl_df$lm_std_beta0_hat
mat_res_fl_df$lower_ci_lm_beta0 = mat_res_fl_df$lm_beta0_hat - zval * mat_res_fl_df$lm_std_beta0_hat
mat_res_fl_df$upper_ci_lm_beta1 = mat_res_fl_df$lm_beta1_hat + zval * mat_res_fl_df$lm_std_beta1_hat
mat_res_fl_df$lower_ci_lm_beta1 = mat_res_fl_df$lm_beta1_hat - zval * mat_res_fl_df$lm_std_beta1_hat
dplyr::between(rep(beta[1], B_fl), mat_res_fl_df$lower_ci_lm_beta0, mat_res_fl_df$upper_ci_lm_beta0) %>% mean()
dplyr::between(rep(beta[2], B_fl), mat_res_fl_df$lower_ci_lm_beta1, mat_res_fl_df$upper_ci_lm_beta1) %>% mean()




# Plot empirical distributions (beta and stochastic parameters)
```

Plot regression estimates for the flicker‑noise case.


```{r}
par(mfrow = c(1, 4))
boxplot_mean_dot(mat_res_df[, c("gmwmx_beta0_hat")],las=1,
                 names = c("beta0"),
                 main = expression(beta[0]), ylab = "Estimate")
abline(h = beta[1], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta1_hat")],las=1,
                 names = c("beta1"),
                 main = expression(beta[1]), ylab = "Estimate")
abline(h = beta[2], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta2_hat")],las=1,
                 names = c("beta2"),
                 main = expression(beta[2]), ylab = "Estimate")
abline(h = beta[3], col = "black", lwd = 2)

boxplot_mean_dot(mat_res_df[, c("gmwmx_beta3_hat")],las=1,
                 names = c("beta3"),
                 main = expression(beta[3]), ylab = "Estimate")
abline(h = beta[4], col = "black", lwd = 2)

par(mfrow = c(1, 1))
```

Plot stochastic parameter estimates for the flicker‑noise case.


```{r}
par(mfrow = c(1, 2))

boxplot_mean_dot(mat_res_fl_df$sigma2_fl,las=1,
                 main = expression(sigma["FL"]^2), ylab = "Estimate")
abline(h = sigma2_fl, col = "black", lwd = 2)

boxplot_mean_dot(mat_res_fl_df$sigma2_wn,las=1,
                 names = c("phi_ar1"),
                 main = expression(sigma["WN"]^2), ylab = "Estimate")
abline(h = sigma2_wn_fl, col = "black", lwd = 2)
par(mfrow = c(1, 1))






```
